### 2022-04-25 paper/tab_nn-architecture.tex

\begin{tabular}{|c|c|c|c|c|c|c|}
    \hline
        variant & $sc_{emb}$      & $sm_{emb}$   & $sh_{emb}$   & $s_{emb}$    & $c_{emb}$    & $fc$ \\
    \hline
                & conv1d(10,20)   & linear(3,64) & linear(1,64) & linear(64,64)& linear(3,64) & linear(64,64)\\
                & relu            & sigmoid      & sigmoid      & sigmoid      & sigmoid      & relu\\
    conv1d-small  & conv1d(20,20)   &              &              &              &              & linear(64,3)\\
                & relu            &              &              &         &         &      \\
                & linear(100,64)  &              &              &         &         &      \\
                & sigmoid         &              &              &         &         &      \\
    \hline
                & conv1d(10,20)   & linear(3,64) & linear(1,64) & linear(64,64)& linear(3,64) & linear(64,64)\\
                & relu            & relu         & relu         & relu         & sigmoid      & relu\\
    conv1d-big & conv1d(20,20)   & linear(64,64)& linear(64,64)&              &              & linear(64,3)\\
                & relu            & sigmoid      & sigmoid      &              &              &      \\
                & linear(100,64)  &          &          &         &         &      \\
                & sigmoid         &          &          &         &         &      \\
    \hline
                & linear(130,64)  & linear(3,64) & linear(1,64) & linear(64,64)& linear(3,64) & linear(64,64)\\
    dense-small & sigmoid         & sigmoid      & sigmoid      & sigmoid      & sigmoid      & relu\\
                &                 &              &              &              &              & linear(64,3)\\
    \hline
                & linear(130,64)  & linear(3,64) & linear(1,64) & linear(64,64)& linear(3,64) & linear(64,64)\\
    dense-big & relu            & relu         & relu         & relu         & sigmoid      & relu\\
                & linear(64,64)   & linear(64,64)& linear(64,64)&              &              & linear(64,3)\\
                & sigmoid         & sigmoid      & sigmoid      &              &              &      \\
    \hline
\end{tabular}