### 2022-07-27 paper/fig_ari.tex

\begin{tikzpicture}
    \pgfplotsset{
      scale only axis,
      height=6cm,
      xlabel={total number of infections},
      ylabel={total lost contacts}
    }
  \begin{axis}[
    legend entries={{$\text{PCN}_\text{ARH}$ (ODE)}, {$\text{PCN}_\text{ARI}$ (ODE)}, {$\text{PCN}_\text{ARH}$ (Binomial)}, {$\text{PCN}_\text{ARI}$ (Binomial)}},
    %, {Fixed}},
    width=9.5cm
  ]
  \addplot[only marks,draw opacity=0.5,mark=x,red] table [x=o_0, y=o_5, col sep=comma] {data/cs/arh/cs_ode.csv};
  \addplot[only marks,draw opacity=0.5,mark=o,blue] table [x=o_0, y=o_5, col sep=comma] {data/cs/ari/cs_ode.csv};
  \addplot[only marks,draw opacity=0.5,mark=triangle,orange] table [x=o_0, y=o_5, col sep=comma] {data/cs/arh/cs_binomial.csv};
  \addplot[only marks,draw opacity=0.5,mark=diamond,cyan] table [x=o_0, y=o_5, col sep=comma] {data/cs/ari/cs_binomial.csv};
  %\addplot[only marks,mark=+] table [x=o_0, y=o_1, col sep=comma] {data/cs/cs_fixed.csv};
  \end{axis}
\end{tikzpicture}