### 2022-07-27 paper/fig_policy-execution.tex

\begin{tikzpicture}
\pgfplotsset{
      scale only axis,
      date coordinates in=x,
      xticklabel=\day/\month,
      height=2cm,
    }
    \begin{groupplot}[
    group style={
            % set how the plots should be organized
            group size=1 by 3,
            % only show ticklabels and axis labels on the bottom
            y descriptions at=edge left,
            x descriptions at=edge bottom,
            % set the `vertical sep' to zero
            horizontal sep=0pt,
            vertical sep=0pt
        },
      width=9.5cm,
      axis y line*=left,
      xlabel={time},
      ylabel={people},
    ]
    % GROUP WITH LEFT Y-AXIS: PEOPLE (HOSP, ICU, DEATHS)
    \nextgroupplot[scaled y ticks=false]
    \addplot[mark=none, cyan] table [x=dates, y=i_hosp_new, col sep=comma] {data/ode/policy_0.csv};
    \label{i_hosp_new}
    \addplot[mark=none, orange] table [x=dates, y=i_icu_new, col sep=comma] {data/ode/policy_0.csv};
    \label{i_icu_new}
    \addplot[mark=none, red] table [x=dates, y=d_new, col sep=comma] {data/ode/policy_0.csv};
    \label{d_new}
    \nextgroupplot[scaled y ticks=false]
    \addplot[mark=none, cyan] table [x=dates, y=i_hosp_new, col sep=comma] {data/ode/policy_30.csv};
    \addplot[mark=none, orange] table [x=dates, y=i_icu_new, col sep=comma] {data/ode/policy_30.csv};
    \addplot[mark=none, red] table [x=dates, y=d_new, col sep=comma] {data/ode/policy_30.csv};
    \nextgroupplot[scaled y ticks=false]
    \addplot[mark=none, cyan] table [x=dates, y=i_hosp_new, col sep=comma] {data/ode/policy_139.csv};
    \addplot[mark=none, orange] table [x=dates, y=i_icu_new, col sep=comma] {data/ode/policy_139.csv};
    \addplot[mark=none, red] table [x=dates, y=d_new, col sep=comma] {data/ode/policy_139.csv};
    \end{groupplot}
    % GROUP WITH RIGHT Y-AXIS: ACTIONS (P_W, P_S, P_L)
    \begin{groupplot}[
    group style={
            % set how the plots should be organized
            group size=1 by 3,
            % only show ticklabels and axis labels on the bottom
            y descriptions at=edge right,
            x descriptions at=edge bottom,
            % set the `vertical sep' to zero
            horizontal sep=0pt,
            vertical sep=0pt
        },
      width=9.5cm,
      axis y line*=right,
      axis x line=none,
      ymax=1.1,
      ylabel={proportion},
      legend style={at={(0.5,1)},anchor=south},
    %   legend pos=outer north east,
      legend columns=3, 
      legend cell align=left,
    ]
    \nextgroupplot[scaled y ticks=false]
    \addplot[mark=none, dashed, blue] table [x=dates, y=p_w, col sep=comma] {data/ode/policy_0.csv};
    \label{p_w}
    \addplot[mark=none, dashed, magenta] table [x=dates, y=p_s, col sep=comma] {data/ode/policy_0.csv};
    \label{p_s}
    \addplot[mark=none, dashed, darkgray] table [x=dates, y=p_l, col sep=comma] {data/ode/policy_0.csv};
    \label{p_l}
    \addlegendimage{/pgfplots/refstyle=i_hosp_new}\addlegendentry{$p_w$}
    \addlegendimage{/pgfplots/refstyle=i_icu_new}\addlegendentry{$p_l$}
    \addlegendimage{/pgfplots/refstyle=d_new}\addlegendentry{$p_s$}
    \addlegendimage{/pgfplots/refstyle=p_w}\addlegendentry{{daily new hosp}}
    \addlegendimage{/pgfplots/refstyle=p_s}\addlegendentry{{daily new ICU}}
    \addlegendimage{/pgfplots/refstyle=p_l}\addlegendentry{{daily deaths}}
    \nextgroupplot[scaled y ticks=false]
    \addplot[mark=none, dashed, blue] table [x=dates, y=p_w, col sep=comma] {data/ode/policy_30.csv};
    \addplot[mark=none, dashed, magenta] table [x=dates, y=p_s, col sep=comma] {data/ode/policy_30.csv};
    \addplot[mark=none, dashed, darkgray] table [x=dates, y=p_l, col sep=comma] {data/ode/policy_30.csv};
    \nextgroupplot[scaled y ticks=false]
    \addplot[mark=none, dashed, blue] table [x=dates, y=p_w, col sep=comma] {data/ode/policy_139.csv};
    \addplot[mark=none, dashed, magenta] table [x=dates, y=p_s, col sep=comma] {data/ode/policy_139.csv};
    \addplot[mark=none, dashed, darkgray] table [x=dates, y=p_l, col sep=comma] {data/ode/policy_139.csv};
    \end{groupplot}
  \end{tikzpicture}