### 2023-01-19 paper/fig_budget_binomial.tex

\begin{tikzpicture}
    \pgfplotsset{
        scale only axis,
        height=6cm,
        xlabel={total number of daily new hospitalizations},
        ylabel={total lost contacts}
    }
    \begin{axis}[
    legend entries={{$\text{PCN}_\text{ARH}$ (Binomial)}, {Budget 2}, {Budget 3}, {Budget 4}, {Budget 5}, {Budget None}},
    width=9.5cm
    ]
    \addplot[only marks,draw opacity=0.5,mark=x,red] table [x=o_1, y=o_5, col sep=comma] {data/cs/arh/cs_binomial.csv};
    \addplot[only marks,draw opacity=0.5,mark=o,blue] table [x=o_1, y=o_5, col sep=comma] {data/cs/budget/binomial/budget_2.csv};
    \addplot[only marks,draw opacity=0.5,mark=square,pink] table [x=o_1, y=o_5, col sep=comma] {data/cs/budget/binomial/budget_3.csv};
    \addplot[only marks,draw opacity=0.5,mark=triangle,orange] table [x=o_1, y=o_5, col sep=comma] {data/cs/budget/binomial/budget_4.csv};
    \addplot[only marks,draw opacity=0.5,mark=diamond,cyan] table [x=o_1, y=o_5, col sep=comma] {data/cs/budget/binomial/budget_5.csv};
    \addplot[only marks,mark=+] table [x=o_1, y=o_5, col sep=comma] {data/cs/budget/binomial/budget_None.csv};
    \end{axis}
\end{tikzpicture}