### 2023-01-31 paper/fig/arh.tex
plot for different budgets

\begin{tikzpicture}
\pgfplotsset{
    scale only axis,
    % height=6cm,
    xlabel={total daily new hospitalizations},
    ylabel={total lost contacts},
    width=0.45\textwidth
}
\begin{groupplot}[
    group style={
        group size=2 by 1,
        y descriptions at=edge left,
        vertical sep=0pt,
        horizontal sep=0pt,
    },
    legend entries={{$\text{PCN}_{b=2}$}, {$\text{PCN}_{b=3}$}, {$\text{PCN}_{b=4}$}, {$\text{PCN}_{b=5}$}, {$\text{PCN}$}, {Fixed}},
]
\nextgroupplot[mark size=1pt,draw opacity=1]
        \addplot[only marks,mark=o,pc1] table [x=o_1, y=o_5, col sep=comma] {data/cs/arh/ode/budget_2.csv};
        \addplot[only marks,mark=o,pc2] table [x=o_1, y=o_5, col sep=comma] {data/cs/arh/ode/budget_3.csv};
        \addplot[only marks,mark=o,pc3] table [x=o_1, y=o_5, col sep=comma] {data/cs/arh/ode/budget_4.csv};
        \addplot[only marks,mark=o,pc4] table [x=o_1, y=o_5, col sep=comma] {data/cs/arh/ode/budget_5.csv};
        \addplot[only marks,mark=o,pc5] table [x=o_1, y=o_5, col sep=comma] {data/cs/arh/ode/budget_None.csv};
        \addplot[only marks,mark=+] table [x=o_0, y=o_1, col sep=comma] {data/cs/cs_fixed.csv};
\nextgroupplot[mark size=1pt]
        \addplot[only marks,draw opacity=1,fill opacity=0.5,mark=o,pc1] table [x=o_1, y=o_5, col sep=comma] {data/cs/arh/binomial/budget_2.csv};
        \addplot[only marks,draw opacity=1,fill opacity=0.5,mark=o,pc2] table [x=o_1, y=o_5, col sep=comma] {data/cs/arh/binomial/budget_3.csv};
        \addplot[only marks,draw opacity=1,fill opacity=0.5,mark=o,pc3] table [x=o_1, y=o_5, col sep=comma] {data/cs/arh/binomial/budget_4.csv};
        \addplot[only marks,draw opacity=1,fill opacity=0.5,mark=o,pc4] table [x=o_1, y=o_5, col sep=comma] {data/cs/arh/binomial/budget_5.csv};
        \addplot[only marks,draw opacity=1,fill opacity=0.5,mark=o,pc5] table [x=o_1, y=o_5, col sep=comma] {data/cs/arh/binomial/budget_None.csv};
        \addplot[only marks,mark=+] table [x=o_0, y=o_1, col sep=comma] {data/cs/cs_fixed.csv};

\end{groupplot}
\end{tikzpicture}