### 2023-01-31 paper/fig/budget_5.tex

\begin{tikzpicture}
    \pgfplotsset{
          scale only axis,
          date coordinates in=x,
          xticklabel=\day/\month,
          height=2cm,
        }
        \begin{groupplot}[
        group style={
                % set how the plots should be organized
                group size=1 by 3,
                % only show ticklabels and axis labels on the bottom
                y descriptions at=edge left,
                x descriptions at=edge bottom,
                % set the `vertical sep' to zero
                horizontal sep=0pt,
                vertical sep=0pt
            },
          width=9.5cm,
          axis y line*=left,
          xlabel={time},
          ylabel={people},
        ]
        % GROUP WITH LEFT Y-AXIS: PEOPLE (HOSP, ICU, DEATHS)
        \nextgroupplot[scaled y ticks=false]
        \addplot[mark=none, cyan] table [x=dates, y=i_hosp_new, col sep=comma] {data/transitions/arh/binomial/budget_5/policies-transitions/0/run_0.csv};
        \label{i_hosp_new}
        \addplot[mark=none, orange] table [x=dates, y=i_icu_new, col sep=comma] {data/transitions/arh/binomial/budget_5/policies-transitions//0/run_0.csv};
        \label{i_icu_new}
        \addplot[mark=none, red] table [x=dates, y=d_new, col sep=comma] {data/transitions/arh/binomial/budget_5/policies-transitions//0/run_0.csv};
        \label{d_new}
        \nextgroupplot[scaled y ticks=false]
        \addplot[mark=none, cyan] table [x=dates, y=i_hosp_new, col sep=comma] {data/transitions/arh/binomial/budget_5/policies-transitions/45/run_0.csv};
        \addplot[mark=none, orange] table [x=dates, y=i_icu_new, col sep=comma] {data/transitions/arh/binomial/budget_5/policies-transitions/45/run_0.csv};
        \addplot[mark=none, red] table [x=dates, y=d_new, col sep=comma] {data/transitions/arh/binomial/budget_5/policies-transitions/45/run_0.csv};
        \nextgroupplot[scaled y ticks=false]
        \addplot[mark=none, cyan] table [x=dates, y=i_hosp_new, col sep=comma] {data/transitions/arh/binomial/budget_5/policies-transitions/114/run_0.csv};
        \addplot[mark=none, orange] table [x=dates, y=i_icu_new, col sep=comma] {data/transitions/arh/binomial/budget_5/policies-transitions/114/run_0.csv};
        \addplot[mark=none, red] table [x=dates, y=d_new, col sep=comma] {data/transitions/arh/binomial/budget_5/policies-transitions/114/run_0.csv};
        \end{groupplot}
        % GROUP WITH RIGHT Y-AXIS: ACTIONS (P_W, P_S, P_L)
        \begin{groupplot}[
        group style={
                % set how the plots should be organized
                group size=1 by 3,
                % only show ticklabels and axis labels on the bottom
                y descriptions at=edge right,
                x descriptions at=edge bottom,
                % set the `vertical sep' to zero
                horizontal sep=0pt,
                vertical sep=0pt
            },
          width=9.5cm,
          axis y line*=right,
          axis x line=none,
          ymax=1.1,
          ylabel={proportion},
          legend style={at={(0.5,1)},anchor=south},
        %   legend pos=outer north east,
          legend columns=3, 
          legend cell align=left,
        ]
        \nextgroupplot[scaled y ticks=false]
        \addplot[mark=none, dashed, blue] table [x=dates, y=p_w, col sep=comma] {data/transitions/arh/binomial/budget_5/policies-transitions//0/run_0.csv};
        \label{p_w}
        \addplot[mark=none, dashed, magenta] table [x=dates, y=p_s, col sep=comma] {data/transitions/arh/binomial/budget_5/policies-transitions//0/run_0.csv};
        \label{p_s}
        \addplot[mark=none, dashed, darkgray] table [x=dates, y=p_l, col sep=comma] {data/transitions/arh/binomial/budget_5/policies-transitions//0/run_0.csv};
        \label{p_l}
        \addlegendimage{/pgfplots/refstyle=i_hosp_new}\addlegendentry{$p_w$}
        \addlegendimage{/pgfplots/refstyle=i_icu_new}\addlegendentry{$p_l$}
        \addlegendimage{/pgfplots/refstyle=d_new}\addlegendentry{$p_s$}
        \addlegendimage{/pgfplots/refstyle=p_w}\addlegendentry{{daily new hosp}}
        \addlegendimage{/pgfplots/refstyle=p_s}\addlegendentry{{daily new ICU}}
        \addlegendimage{/pgfplots/refstyle=p_l}\addlegendentry{{daily deaths}}
        \nextgroupplot[scaled y ticks=false]
        \addplot[mark=none, dashed, blue] table [x=dates, y=p_w, col sep=comma] {data/transitions/arh/binomial/budget_5/policies-transitions/45/run_0.csv};
        \addplot[mark=none, dashed, magenta] table [x=dates, y=p_s, col sep=comma] {data/transitions/arh/binomial/budget_5/policies-transitions/45/run_0.csv};
        \addplot[mark=none, dashed, darkgray] table [x=dates, y=p_l, col sep=comma] {data/transitions/arh/binomial/budget_5/policies-transitions/45/run_0.csv};
        \nextgroupplot[scaled y ticks=false]
        \addplot[mark=none, dashed, blue] table [x=dates, y=p_w, col sep=comma] {data/transitions/arh/binomial/budget_5/policies-transitions/114/run_0.csv};
        \addplot[mark=none, dashed, magenta] table [x=dates, y=p_s, col sep=comma] {data/transitions/arh/binomial/budget_5/policies-transitions/114/run_0.csv};
        \addplot[mark=none, dashed, darkgray] table [x=dates, y=p_l, col sep=comma] {data/transitions/arh/binomial/budget_5/policies-transitions/114/run_0.csv};
        \end{groupplot}
      \end{tikzpicture}